---
title: "Data Visualization Project"
author: "Juan David Jaramillo"
format: html
editor: visual
---

# Inequaility Income

Having grown up in one of the most unequal societies in the world (Latin America), I have been able to observe the significant impact that economic inequality has on a society, on the behavior of its individuals, their aspirations, and their ideals. While this is a topic with a deep sociological perspective and extensive complexity, for this project, I would like to address simple and interesting relationships between variables related to economic inequality in different countries around the world.

## Summary of Data Tables

For this project, three different databases will be used. One of them contains information on socioeconomic inequality indices and Human Development by country (obtained from Kraggle) . The second database includes information on the cost of living, purchasing power, and average monthly income by country (obtained from Kraggle). Lastly, the third database provides information on the percentage of total monthly income generated in a country that is obtained by the top 10% richest of the population (obtained from World Inequality Database).

To create the master database, the three databases are merged using the variable "Country" as the key. As a result, we obtain a database with the following information for more than 100 countries:

```{r}
library(data.table)
library(kableExtra)
library(scales)
library(ggplot2)
library(tidyr)
library(dplyr)
library(knitr)
db1<-fread("cost of life.csv")
db2<-fread("inequality in income.csv")
db3<-fread("wealthiest.csv")
data<-merge(db2,db1,by="country",all.x = T)
data<-merge(data,db3,by="country",all.x = T)
vars<-names(data)
descvars<-c("Name of the Country",
            "Name of the Continent",
            "Human Development Group: Low,Medium,High,Very High", "Index of Inequality Income: Number from 0 to 100","","","","","","","","","","","","","Cost of Life Index:  A numerical value that quantifies the cost of living in each country
","Purchasing Power Index: A numerical value that quantifies the purchasing power of the population in each country","Percentage of the total income of a country which is obtained for the 10% wealthiest")
summaryvars<-data.frame("Variables"=vars,"Description"=descvars)
kable(summaryvars,format = "html") |> kable_styling(font_size = 12)
```

## Income Inequality

To begin with, it is intriguing to examine the behavior of **InequalityIncome** over recent years. To this end, curves can be generated to illustrate how this variable has evolved overall, by continents, and by HDI groups.

```{r}
 data_clean <- data[complete.cases(data[, .(InequalityIncome2010, InequalityIncome2011, InequalityIncome2012, 
                                           InequalityIncome2013, InequalityIncome2014, InequalityIncome2015,
                                           InequalityIncome2016, InequalityIncome2017, InequalityIncome2018,
                                           InequalityIncome2019, InequalityIncome2020, InequalityIncome2021)]), ]
data_long <- data_clean |> 
  gather(key = "Year", value = "InequalityIncome", 
         InequalityIncome2010:InequalityIncome2021) |> 
  mutate(Year = as.numeric(gsub("InequalityIncome", "", Year)))
# Avg Per Continent
data_avg <- data_long  |> 
  group_by(Year, continent) |>
  summarize(avg_inequality = mean(InequalityIncome, na.rm = TRUE))

# Global Avg
data_global_avg <- data_long |>
  group_by(Year) |>
  summarize(avg_inequality = mean(InequalityIncome, na.rm = TRUE))
data_global_avg$continent<-"Worldwide"
data_global_avg<- data_global_avg |> select(-avg_inequality,avg_inequality)
total_avg_data<-rbind(data_avg,data_global_avg)

```

```{r}

ggplot(data= total_avg_data, aes(
    x = Year,
    y = avg_inequality,
    colour = continent,
    group = continent
  )) +
  geom_line(size = 0.8) +
  scale_colour_manual(
    values = c("Worldwide" = "black", 
               "Africa" = "red4", 
               "Asia" = "olivedrab", 
               "Europe" = "cyan4", 
               "Oceania" = "royalblue4", 
               "America" = "gold1")
  ) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "grey", linewidth = 0.5),
    panel.grid.minor = element_line(color = "grey", linewidth = 0.25),
    legend.position = "none",
  ) +
  scale_y_continuous(
        sec.axis = dup_axis(
      breaks = total_avg_data |>
        group_by(continent) |>
        summarise(avg_inequality = last(avg_inequality)) |>
        pull(avg_inequality),
      labels = total_avg_data |>
        group_by(continent) |>
        summarise(avg_inequality = last(avg_inequality)) |>
        pull(continent),
      name = NULL
    ),   
    expand = c(0, 0),
    limits = c(10, max(total_avg_data$avg_inequality) * 1.1)
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = seq(min(total_avg_data$Year), max(total_avg_data$Year), by = 1)
  ) +
  labs(
    title = "Average Inequality Income Index by Continent",
    y = "Inequaility Income Index",
    x = "Year"
  ) +
  theme(axis.title = element_text(size = 9)) +
  geom_text(
    data = total_avg_data |>  
      group_by(continent) |> 
      slice_tail(n = 1),  # Selecciona el último valor de cada continente
    aes(label = continent, colour = continent),
    hjust = -0.2,  # Ajusta la posición de la etiqueta
    size = 3.5
  ) -> IIContinentGraph







```

```{r}

# Avg Per HDGroup
data_avg_HD <- data_long  |> 
  group_by(Year, HDGroup) |>
  summarize(avg_inequality = mean(InequalityIncome, na.rm = TRUE))

# Global Avg
data_global_avg <- data_long |>
  group_by(Year) |>
  summarize(avg_inequality = mean(InequalityIncome, na.rm = TRUE))
data_global_avg$HDGroup<-"Worldwide"
data_global_avg<- data_global_avg |> select(-avg_inequality,avg_inequality)
total_avg_data2<-rbind(data_avg_HD,data_global_avg)

```

```{r}

ggplot(data= total_avg_data2, aes(
    x = Year,
    y = avg_inequality,
    colour = HDGroup,
    group = HDGroup,
  )) +
  geom_line(size = 0.8) +
  scale_colour_manual(
    values = c("Worldwide" = "black", 
               "Low" = "red4", 
               "Medium" = "gold1", 
               "High" = "green4", 
               "Very High" = "dodgerblue3")
  ) +
  theme(
    panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
    panel.background = element_rect(fill = "white"),
    panel.grid.major = element_line(color = "grey", linewidth = 0.5),
    panel.grid.minor = element_line(color = "grey", linewidth = 0.25),
    legend.position = "none",
  ) +
  scale_y_continuous(
        sec.axis = dup_axis(
      breaks = total_avg_data2 |>
        group_by(HDGroup) |>
        summarise(avg_inequality = last(avg_inequality)) |>
        pull(avg_inequality),
      labels = total_avg_data2 |>
        group_by(HDGroup) |>
        summarise(avg_inequality = last(avg_inequality)) |>
        pull(HDGroup),
      name = NULL
    ),   
    expand = c(0, 0),
    limits = c(10, max(total_avg_data2$avg_inequality) * 1.1)
  ) +
  scale_x_continuous(
    expand = c(0, 0),
    breaks = seq(min(total_avg_data2$Year), max(total_avg_data2$Year), by = 1)
  ) +
  labs(
    title = "Average Inequality Income Index by Human Development Group",
    y = "Inequaility Income Index",
    x = "Year"
  ) +
  theme(axis.title = element_text(size = 9)) +
  
  geom_text(
    data = total_avg_data2 |>  
      group_by(HDGroup) |> 
      slice_tail(n = 1),
    aes(label = HDGroup, colour = HDGroup),
    hjust = -0.2, 
    size = 3.5
  ) -> IIHDGraph



```

```{r,dpi=900}

print(IIContinentGraph)
print(IIHDGraph)
```

### Average inequality income by Continent

In the first graph, we can observe that in the last decade, the general trend of inequality is decreasing, except for Oceania, which seems to maintain stable values.

If we analyze the behavior by continent, we can note that America has had the most variation in the last decade, lowering its average Inequality Income by almost 3 points.

Regarding Europe, we can note that between 2010 and 2016, it was the only continent that showed slight constant growth in inequality. However, from 2016 to 2021, it was able to recover significantly.

Overall, we see that Africa and America are far above the rest of the continents in terms of inequality levels.

### Average inequality income by Human Development Group

On the other hand, in the second graph, surprisingly, we find that the Human Development Index is not closely related to the Inequality Income Index.

While countries with a very high Human Development Index are far below the average, the rest of the countries, including those with a High Human Development Index, always exceed 25 points in the inequality index. Furthermore, these countries tend to behave very similarly in the last five years of the study period, regardless of their Human Development Group.

## Cost of Life

Continuing with the analysis, it is worth exploring how countries are distributed in terms of the cost of living and purchasing power. Additionally, we can relate these variables to the HDGroup for a sample composed of countries that are interesting to our audience!

```{r}
filtered_data <- data |> 
  filter(!is.na(cost_index), !is.na(purchasingpower_index))

# Crear el gráfico de dispersión
ggplot(data = filtered_data, aes(x = cost_index, y = purchasingpower_index, colour = continent)) +
    scale_colour_manual(
    values = c("Worldwide" = "black", 
               "Africa" = "red4", 
               "Asia" = "olivedrab", 
               "Europe" = "cyan4", 
               "Oceania" = "royalblue4", 
               "America" = "gold1")
  ) +
  geom_point(size= 2.5) +
    geom_abline(slope = 1, intercept = 0, color = "black", linetype = "dashed", size = 1) + 
  labs(
    title = "Cost of Life vs Power Purchasing",
    caption = "Information available for 98 countries",
    x = "Cost of Life",
    y = "Power Purchasing",
    color = "Continent"
  ) +
  theme(plot.subtitle = element_text(size = 13),
        axis.title   = element_text(size = 11)) +
  theme(legend.position = "bottom",
        panel.border = element_rect(colour = "black", fill = NA, size = 0.8),
       panel.background = element_rect(fill = "white"),
       panel.grid.major = element_line(color = "grey", linewidth = 0.5),
       panel.grid.minor = element_line(color = "grey", linewidth = 0.25),
        legend.background = element_rect(fill = "white", color = NA)) 

```
